Classless Java: Tuning Java Interfaces

### 2015-12-03 paper/sections/Appendix.tex
Big progress on formalism/proofs

\textbf{LEMMA 1. }
For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} and $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
\begin{proof}
By induction on the typing rules: by the grammar shown in Figure~\ref{Grammar}, there are 6 cases for an arbitrary expression $e$:
\begin{itemize}
\item Variables are typed in the same exact way.
\item Field update. The type preservation is ensured by induction.
\item A method call (normal, static or super). The corresponding method declaration won't be ``removed'' by the translation, also the types remain unchanged. The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
\item An object creation. Adding the $\QM{of}$ method doesn't introduce unimplemented methods to an interface, moreover, the static method is not inheritable, hence after translation such an object creation still type checks and has the right type by induction.
\end{itemize}
\end{proof}

\subsection{LEMMA 2 and Proof}\label{subsec:lemma2}

\textbf{LEMMA 2. }
If $\II_0$ has \textbf{@ObjWeak} and $\C_0$ OK, then $[\![\II_0]\!]$ OK.
\begin{proof}

@@

\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\II_0]\!]$, the type of the return value is compatible with the method's return type.

Since $\II_0$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.

By the definition of $\ofMethod(\C_0)$, the return value is an object $$\QM{return new}\ \C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
To prove it is of type $\C_0$, we use the typing rule \rn{t-Obj}.

@@

\item All field initializations are type correct. By the definition of $\ofMethod(\C_0)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.

@@

    \item Typing of the $i$-th getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]

@@

    \item Typing of the \QM{with-} method of an arbitrary field $m_i$. By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C_0\QM{.of(}\es_i\QM{);\}}\]
        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C_0.\QM{of}$ method. Hence \[\Gamma, m_1:\C_1\ldots m_n:\C_n, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.

@@

        \[\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
        By \rn{t-Update}, the assignment ``$\m_i\QM{= _val;}$'' is correct since $\m_i$ and $\QM{_val}$ have the same type $I_i$, and the return type is decided by \QM{this}. \[\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash\QM{this}\in \C_0\]
        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
        \haoyuan{Can we write $\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i, \C^{\mh_i} = \C_0 \vdash\QM{this}\in \C_0 = \C^{\mh_i}$?}\\
        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C_0$?}

@@

\item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
    For convenience, we use ``$\method$ in $\ofMethod(\C_0)$'' to denote that $\method$ is one of the implemented methods in the return expression of $\ofMethod(\C_0)$, namely \Q@new@ $\C_0$\Q@(){...}@.

@@

        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C_0)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; = \mBody(m_i,\C_0)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C_0)
        \end{align*}\marco{I do not understand this point, I think ofMethod is not what you want to write, but I can not understand what you want do do...}
        \haoyuan{Right, sorry for that, I mean if a getter is created by ofMethod, then it is a field of I0, and finally for this part our aim is to prove sigvalid, and the definition of sigvalid asks you to prove $\mh\QM;<:\mBody(...)$, see the last line.}

@@

        &\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
        \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
        \mimply\hspace{.2in}& \isWith(\mBody(\QM{with#}\m_i,\C_0),\C_0)\\
        \mimply\hspace{.2in}& \C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM{with#}\m_i,\C_0)

@@

        &\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
        \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
        \mimply\hspace{.2in}& \isSetter(\mBody(\QM_\m_i,\C_0),\C_0)\\
        \mimply\hspace{.2in}& \C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM_\m_i,\C_0)

@@

\item All abstract methods in the domain of $\C_0$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
    Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases. A getter $m_i$ is generated if it satisfies $\isField$; a \Q@with-@ method is generated for the case $\isWith$, by the definition of $\withMethod$; a setter for $\isSetter$, similarly, by the definition of $\setterMethod$. Hence it is of type $\C_0$ by \rn{t-Obj}.

@@

\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$
Since $\II_0$ OK, we have $\dom(\II_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\method')$$
This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so $\method'$ doesn't overlap with $\dom(\C_0)$. The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$. Also note that after translation, there are still no methods with conflicted names, since the \QM{of} method was previously not in the domain, hence $[\![\II_0]\!]$ is well-formed, which finishes our proof.
\end{proof}

\subsection{THEOREM and Proof}\label{subsec:theorem}
\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
if $\II_0$ has \textbf{@ObjWeak},
$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$\\*
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.
\begin{proof}
\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, for any $\II\ne\II_0$, by \textbf{LEMMA 1}, we know that all its methods
are still well-typed, and the generated code in translation of \textbf{@ObjWeak} is only a static method \QM{of}, which has no way to affect the domain
of $\II$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
\end{proof}

>>>>

\textbf{LEMMA 1. }
For any expression $e$ under an interface table $\II$ IT where $\Gamma\vdash e\in\C^\II$, $\II$ has \textbf{@ObjWeak} annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\C^\II$.
\begin{proof}
By induction on the typing rules: by the grammar shown in Figure~\ref{Grammar}, there are 6 cases for an arbitrary expression $e$:
\begin{itemize}
\item Variables are typed in the same exact way.
\item Field update. The type preservation is ensured by induction.
\item A method call (normal, static or super). The corresponding method declaration won't be ``removed'' by the translation, also the types remain unchanged. The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
\item An object creation. Adding the $\QM{of}$ method doesn't introduce unimplemented methods to an interface, moreover, the static method is not inheritable, hence after translation such an object creation still type checks and has the right type by induction.
\end{itemize}
\end{proof}
\textbf{LEMMA 1b. }
For any expression $e$ under an interface table $\II$ IT where there is no heir of $\C^\II$,  $\Gamma\vdash e\in\C^\II$, $\II$ has \mixin annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\_<:\C^\II$.
\begin{proof}
The proof follows the same scheme of the Lemma1, but for the case of method call the return type may be refined with a subtype. This is still ok since we require $\_<:\C^\II$. On the other side, this weaker result still allows the application on the method call typing rules, since in the premises the types of the actual parameter are required to be a subtype of the formal one.
\end{proof}
\subsection{LEMMA 2 and Proof}\label{subsec:lemma2}

\textbf{LEMMA 2. }
If $\II$ has \textbf{@ObjWeak} annotation and $\C^{\II}$ OK in $\II$ IT, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
\begin{proof}

@@

\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\C^\II]\!]$, the type of the return value is compatible with the method's return type.

Since $\II$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C)$.

By the definition of $\ofMethod(\C)$, the return value is an object $$\QM{return new}\ \C^\II \oR\cR\ \QM{\{}...\ \QM{\}}$$
To prove it is of type $\C^\II$, we use the typing rule \rn{t-Obj}.

@@

\item All field initializations are type correct. By the definition of $\ofMethod(\C^\II)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.

@@

    \item Typing of the $i$-th getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C^\II \vdash m_i\in \C_i\]

@@

    \item Typing of the \QM{with-} method of an arbitrary field $m_i$. By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C^\II\QM{.of(}\es_i\QM{);\}}\]
        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C^\II.\QM{of}$ method. Hence \[\Gamma, m_1:\C_1\ldots m_n:\C_n, \QM{this}:\C^\II,\ \QM{_val} : \C_i \vdash \C^\II\QM{.of}\oR\es_i\cR\in \C^\II\]
        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.

@@

        \[\C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
        By \rn{t-Update}, the assignment ``$\m_i\QM{= _val;}$'' is correct since $\m_i$ and $\QM{_val}$ have the same type $I_i$, and the return type is decided by \QM{this}. \[\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i \vdash\QM{this}\in \C^\II\]
        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
        \haoyuan{Can we write $\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i, \C^{\mh_i} = \C^\II \vdash\QM{this}\in \C^\II = \C^{\mh_i}$?}\\
        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C^\II$?}

@@

\item All method headers are valid with respect to the domain of $\C^\II$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
    For convenience, we use ``$\method$ in $\ofMethod(\C^\II)$'' to denote that $\method$ is one of the implemented methods in the return expression of $\ofMethod(\C^\II)$, namely \Q@new@ $\C^\II$\Q@(){...}@.

@@

        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C^\II)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; = \mBody(m_i,\C^\II)\\
        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C^\II)

@@

        &\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
        \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C^\II) \mbox{ is of form }\mh\QM;\\
        \mbox{with}\hspace{.2in}& \valid(\C^\II)\\
        \mimply\hspace{.2in}& \isWith(\mBody(\QM{with#}\m_i,\C^\II),\C^\II)\\
        \mimply\hspace{.2in}& \C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM{with#}\m_i,\C^\II)

@@

        &\C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
        \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C^\II) \mbox{ is of form }\mh\QM;\\
        \mbox{with}\hspace{.2in}& \valid(\C^\II)\\
        \mimply\hspace{.2in}& \isSetter(\mBody(\QM_\m_i,\C^\II),\C^\II)\\
        \mimply\hspace{.2in}& \C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM_\m_i,\C^\II)

@@

\item All abstract methods in the domain of $\C^\II$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
    Here we simply refer to $\valid(\C^\II)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases. A getter $m_i$ is generated if it satisfies $\isField$; a \Q@with-@ method is generated for the case $\isWith$, by the definition of $\withMethod$; a setter for $\isSetter$, similarly, by the definition of $\setterMethod$. Hence it is of type $\C^\II$ by \rn{t-Obj}.

@@

\noindent\textbf{Part II.} Next we check that in $[\![\II]\!]$, $$\dom([\![\II]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$

Since $\II$ OK, we have $\dom(\II)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II]\!])=\dom(\C^\II)\cup\dom(\method')$$
This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so $\method'$ doesn't overlap with $\dom(\C^\II)$. The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II]\!])$ is only an extension to $\dom(\C)$ with the \QM{of} method, namely $\method'$. Also note that after translation, there are still no methods with conflicted names, since the \QM{of} method was previously not in the domain, hence $[\![\II]\!]$ is well-formed, which finishes our proof.
\end{proof}

\textbf{LEMMA 2b. }
If $\II$ has \mixin annotation $\C^{\II}$ OK in $\II$ IT
and there is no heir of $\C^\II$, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
\begin{proof}
\noindent\textbf{Part I.} Similarly to what already argued for Lemma2, 
since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$,IT.
The translation function delegate its work to \Q!@ObjWeak! in such way that we can refer to 
Lemma2  to complete this part. Note that all the methods that are added (directly) by \mixin are abstract, and thus there is no body to typecheck.

\noindent\textbf{Part II.}
Similar to what already argued for Lemma2, but we need to notice that the the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
Thus by last clause of the definition of $\override(\_)$, $\mBody(\_)$ is defined on the same method names.
\end{proof}

\subsection{THEOREM and Proof}\label{subsec:theorem}

\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
If a given interface table $\II$ IT is OK\\*
 where $\II$ has \Q!@ObjWeak!, 
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
then the interface table $[\![\II]\!]$ IT is OK.

\begin{proof}
\textbf{LEMMA 2} already proves that $[\![\II]\!]$ is OK. On the other hand, for any $\II'\in $ IT$\setminus\II$, by \textbf{LEMMA 1}, we know that all its methods
are still well-typed, and the generated code in translation of \textbf{@ObjWeak} is only a static method \QM{of}, which has no way to affect the domain
of $\II'$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
\end{proof}

\noindent\textbf{THEOREM }\Q!@Obj!\\*
If a given interface table $\II$ IT is OK\\*
 where $\II$ has \Q!@Obj!, 
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
then the interface table $[\![\II]\!]$ IT is OK.

\begin{proof}
Similar to what already argued for \textbf{THEOREM }\Q!@ObjWeak!
we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finsh by \textbf{THEOREM }\Q!@ObjWeak!.
\end{proof}

### 2015-12-03 paper/sections/Translation.tex
Big progress on formalism/proofs

The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$, \hl{since overloading is not modelled}.

@@

\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
That is, the annotation do not remove any functionality, is just
adding more behaviour.
\mixin  does not guarantee \textit{Client coherence}.

@@

\noindent By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
This would break \Q@C@. And it's similar why client coherence is not supported as well. In the client code, an expression
that uses ``\Q@new B(){...}@'' to create an object of \Q@B@ gives an implementation of \Q@A withX(int x);@, but after the return
type is refined to \Q@B@, there is a type error.
\end{itemize}

\noindent Consequently, type refinement is a common case in client code, and hence is included by \mixin as a feature for convenient use, whereas it simultaneously brings some risk in throwing type errors, though they are likely to be fixed easily by users. On the other hand, a property of \mixin is informally concluded as: for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation. Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence. The formal theorem and proof is presented afterwards.

\subsection{Results}

\haoyuan{We use \textbf{@ObjWeak} or \Q:@ObjWeak:?}

The theorem is presented below. Lemmas and the complete proof are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\

\noindent\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
if $\II_0$ has \textbf{@ObjWeak},
$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$\\*
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.

%The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed. In Java it implies not touching the annotated interface, and surely an alternative approach is to create a new interface and generate what we need. But since we are using Lombok for the translation, the transformation is applied to the annotated interface itself. But heir coherence is indeed an issue. See below how this is affected by our approach:

%Now we go back and read the statement of our theorem again. It illustrates that a program with the translation of \mixin applied still type checks if it did, except for the subtypes of the annotated interface (type safety of them is not guaranteed).

%\haoyuan{Text needs improving.}

%\subsubsection{Derived notations}
% Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
%\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]

%\marco{we also need to define a function that gives all the methods of an interface, something line}
%\[
%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
%\]

>>>>

The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only in order to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$. \marco{no, overloading is not the issue here\hl{since overloading is not modelled}}.

@@

\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
That is, the annotation do not remove any functionality, is just
adding more behaviour.
We guarantee \textit{Client coherence}.

@@

\noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
This would break \Q@C@. 
Similarly, an expression of the form \Q@new B(){.. A withX(int x){..}}@ would compile before translation, but would be ill typed after the translation.
\end{itemize}

\noindent Our automatic type refinement 
is a usefull and convenient feature, but is not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantic  to be well typed.

\subsection{Results}
To formally characterize the behaviour of our annotation, and the two level of guarancees that we offer, we provide some notations and two theorems:
\begin{itemize}
\item We say that an interface table 
IT is OK if under such interface table, all interfaces are OK.
\item Since interface tables are just represented as sequences of interfaces we can write IT = $\II$ IT' to talk about a specific interface in a table.
\item We denote with $\C^{\II}$ the name of an interface.
\item An interface table contains a hair of $\C$ if there is any interface that extends it, or any \Q@new@ that instantiate it.
\end{itemize}

\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
If a given interface table $\II$ IT is OK\\*
 where $\II$ has \Q!@ObjWeak!, 
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
then the interface table $[\![\II]\!]$ IT is OK.

\noindent\textbf{THEOREM }\Q!@Obj!\\*
If a given interface table $\II$ IT is OK\\*
 where $\II$ has \Q!@Obj!, 
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
then the interface table $[\![\II]\!]$ IT is OK.

Informally, this means that for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation. Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence. The formal theorems proofs are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
